---
title: "R Notebook"
output: html_notebook
---

####NSDUH 2015 - 2019 Psychological Distress, Suicidality, and Major Depressive Episodes  + MDMA/ecstasy and Psilocybin analyses######
##### Maya A. Seale ################
####### May 11, 2026##################



The goal of this project is to assess sexual orientation as a moderator between the associations of lifetime use of each of two classic psychedelic substances, Psilocybin and MDMA, and psychopathological outcomes of suicidality, psychological distress, and major depressive episodes
I hypothesized that sexual orientation would moderate this association

The full manuscript detailing background, methods, and discussion  can be found here: N/A not yet published

The three sexual orientation groups investigated are "Heterosexual" "Bisexual" and"Gay/lesbian" as constrained by the NSDUH questionnaire

## Reproducibility notes

Data: NSDUH public-use files, 2015–2019 (SAMHSA; available from the Substance Abuse and Mental Health Data Archive, https://www.datafiles.samhsa.gov/). Place the annual files (NSDUH2015Data.RDS … NSDUH2019Data.RDS) in the working directory before running.

This document runs in R; R and package versions are recorded by `sessionInfo()` in the final chunk. No random-number generation is used, so results are deterministic.

# Importing Libraries

```{r}
# Importing Libraries 
if (!require("pacman")) {install.packages("pacman"); require("pacman")}

p_load("survey", "plyr", "labelled", "knitr", "dplyr", "tableone")

```

# Helper functions: consistent ORs, CIs, and p-values for svyglm
```{r}

# CI on the log scale (drop-in replacement): exp(cbind(OR=coef(m), or_confint(m)))
or_confint <- function(model, level = 0.95) {
  co  <- summary(model)$coefficients
  est <- co[, "Estimate"]; se <- co[, "Std. Error"]
  tcrit <- qt(1 - (1 - level)/2, model$df.residual)
  ci <- cbind(est - tcrit*se, est + tcrit*se)
  colnames(ci) <- c("2.5 %", "97.5 %"); rownames(ci) <- rownames(co); ci
}

# Tidy OR / 95% CI / p table (CI and p guaranteed consistent)
or_table <- function(model, level = 0.95, digits = 2) {
  co  <- summary(model)$coefficients
  est <- co[, "Estimate"]; se <- co[, "Std. Error"]
  tcrit <- qt(1 - (1 - level)/2, model$df.residual)
  data.frame(term = rownames(co),
             OR    = round(exp(est), digits),
             lower = round(exp(est - tcrit*se), digits),
             upper = round(exp(est + tcrit*se), digits),
             p     = round(co[, "Pr(>|t|)"], 5), row.names = NULL)
}

# Flags any term where CI and p still disagree (should print "OK")
check_consistency <- function(model) {
  t <- or_table(model)
  excl1 <- (t$lower > 1 & t$upper > 1) | (t$lower < 1 & t$upper < 1)
  bad <- t[excl1 != (t$p < .05), ]
  if (nrow(bad) == 0) message("OK: CIs and p-values agree.") else print(bad)
  invisible(bad)
}
```

# Importing the NSDUH dataset by year
```{r}
PUF2015_021518 <- readRDS("NSDUH2015Data.RDS")
PUF2016_022818 <- readRDS("NSDUH2016Data.RDS")
PUF2017_100918 <- readRDS("NSDUH2017Data.RDS")
PUF2018_100819 <- readRDS("NSDUH2018Data.RDS")
PUF2019_100920 <- readRDS("NSDUH2019Data.RDS")
```

# Selecting the variables needed from each year
```{r}

DATA2015 <- PUF2015_021518 %>% dplyr::select(QUESTID2, 
amdelt, amdeyr, amdeimp,mhsuithk, mhsuipln, spdmon, # Dependent Variables
ecstmoflag,  PSILCY2, # Independent Variables                                     
sexident, # Moderator 
AGE2,irmaritstat, IREDUHIGHST2, CATAG6, irsex, NEWRACE2, income,  # Demographic Covariates 
rskyfqtes, cocflag, herflag, pcpflag, ecstmoflag, inhalflag , pnranyflag, trqanyflag,ketminflag, AMYLNIT2, stmanyflag, sedanyflag, mrjflag,  lsdflag, PEYOTE2, MESC2, # Drug and Behavioral Covariates
ANALWT_C, vestr, verep, # Survey Weights

)

DATA2015 <- DATA2015 %>%
  mutate(year = 2015)
```

```{r}
DATA2016 <- PUF2016_022818 %>% dplyr::select(QUESTID2, 
amdelt, amdeyr, amdeimp,mhsuithk, mhsuipln, spdmon, # Dependent Variables
ecstmoflag,  PSILCY2, # Independent Variables                                     
sexident, # Moderator 
AGE2,irmarit, IREDUHIGHST2, CATAG6, irsex, NEWRACE2, income,  # Demographic Covariates 
rskyfqtes, cocflag, herflag, pcpflag, ecstmoflag, inhalflag , pnranyflag, trqanyflag,ketminflag, AMYLNIT2, stmanyflag, sedanyflag, mrjflag,  lsdflag, PEYOTE2, MESC2, # Drug and Behavioral Covariates
ANALWT_C, vestr, verep, # Survey Weights

)
DATA2016 <- DATA2016 %>%
  mutate(year = 2016)
```

```{r}
DATA2017 <- PUF2017_100918 %>% dplyr::select(QUESTID2, 
amdelt, amdeyr, amdeimp,mhsuithk, mhsuipln, spdmon, # Dependent Variables
ecstmoflag,  PSILCY2, # Independent Variables                                     
sexident, # Moderator 
AGE2,irmarit, IREDUHIGHST2, CATAG6, irsex, NEWRACE2, income,  # Demographic Covariates 
rskyfqtes, cocflag, herflag, pcpflag, ecstmoflag, inhalflag , pnranyflag, trqanyflag,ketminflag, AMYLNIT2, stmanyflag, sedanyflag, mrjflag,  lsdflag, PEYOTE2, MESC2, # Drug and Behavioral Covariates
ANALWT_C, vestr, verep, # Survey Weights
)
DATA2017 <- DATA2017 %>%
  mutate(year = 2017)
```

```{r}
DATA2018 <- PUF2018_100819 %>% dplyr::select(
QUESTID2, 
amdelt, amdeyr, amdeimp,mhsuithk, mhsuipln, spdmon, # Dependent Variables
ecstmoflag,  PSILCY2, # Independent Variables                                     
sexident, # Moderator 
AGE2,irmarit, IREDUHIGHST2, CATAG6, irsex, NEWRACE2, income,  # Demographic Covariates 
rskyfqtes, cocflag, herflag, pcpflag, ecstmoflag, inhalflag , pnranyflag, trqanyflag,ketminflag, AMYLNIT2, stmanyflag, sedanyflag, mrjflag,  lsdflag, PEYOTE2, MESC2, # Drug and Behavioral Covariates
ANALWT_C, vestr, verep, # Survey Weights

)
DATA2018 <- DATA2018 %>%
  mutate(year = 2018)
```

```{r}
DATA2019 <- PUF2019_100920 %>% dplyr::select(
QUESTID2, 
amdelt, amdeyr, amdeimp,mhsuithk, mhsuipln, spdmon, # Dependent Variables
ecstmoflag,  PSILCY2, # Independent Variables                                     
sexident, # Moderator 
AGE2,irmarit, IREDUHIGHST2, CATAG6, irsex, NEWRACE2, income,  # Demographic Covariates 
rskyfqtes, cocflag, herflag, pcpflag, ecstmoflag, inhalflag , pnranyflag, trqanyflag,ketminflag, AMYLNIT2, stmanyflag, sedanyflag, mrjflag,  lsdflag, PEYOTE2, MESC2, # Drug and Behavioral Covariates
ANALWT_C, vestr, verep, # Survey Weights
)
DATA2019 <- DATA2019 %>%
  mutate(year = 2019)
```


## Merge each year into one DataFrame
```{r}

NSDUH2015to2019DataList <- list(DATA2015, DATA2016, DATA2017, DATA2018, DATA2019)                                                                                                                                                  
NSDUH2015to2019Data  <- Reduce(
  function(x, y) merge(x, y, all=TRUE), NSDUH2015to2019DataList, accumulate=FALSE)

##Exclude kids
NSDUH2015to2019Data <-subset(NSDUH2015to2019Data,!CATAG6=="1")

```


### Variable cleaning and collapsing 
#### Create Marriage Variable (combining 2015 with other years)
```{r}
# Create Marriage Variable (combining 2015 with other years)

# Remove irmarit -9 values (equivalent to not present) by pooling with irmaritstat (2015 variable only)
# irmaritstat is 2015 only (they say to consider before pooling, but because it's a covariate, it's okay. under guidance of Grant Jones)
plyr::count(NSDUH2015to2019Data$irmarit)
NSDUH2015to2019Data$MarriageVariable <- ifelse(is.na(NSDUH2015to2019Data$irmarit), NSDUH2015to2019Data$irmaritstat, NSDUH2015to2019Data$irmarit)
plyr::count(NSDUH2015to2019Data$MarriageVariable)
```

#### Create Risk Variable 
```{r}
# Create Risk Variable 
NSDUH2015to2019Data$RiskyVariable <- NSDUH2015to2019Data$rskyfqtes
#Remove all data with a response above 4 (essentially an NA)
NSDUH2015to2019Data$RiskyVariable <- ifelse (NSDUH2015to2019Data$RiskyVariable >4, NA, NSDUH2015to2019Data$RiskyVariable)
plyr::count(NSDUH2015to2019Data$RiskyVariable)
```

#### Recode Outcome Variables to exclude -9 values 
```{r}
###Recode -9 (no data for that year) to NA for the six analyzed outcome variables.
NSDUH2015to2019Data$spdmon <- ifelse(NSDUH2015to2019Data$spdmon == -9, NA, NSDUH2015to2019Data$spdmon)
NSDUH2015to2019Data$mhsuithk <- ifelse(NSDUH2015to2019Data$mhsuithk == -9, NA, NSDUH2015to2019Data$mhsuithk)
NSDUH2015to2019Data$mhsuipln <- ifelse(NSDUH2015to2019Data$mhsuipln == -9, NA, NSDUH2015to2019Data$mhsuipln)
NSDUH2015to2019Data$amdeyr <- ifelse(NSDUH2015to2019Data$amdeyr == -9, NA, NSDUH2015to2019Data$amdeyr)
NSDUH2015to2019Data$amdelt <- ifelse(NSDUH2015to2019Data$amdelt == -9, NA, NSDUH2015to2019Data$amdelt)
NSDUH2015to2019Data$amdeimp <- ifelse(NSDUH2015to2019Data$amdeimp == -9, NA, NSDUH2015to2019Data$amdeimp)
```

#### Code independent variables to be "0" and "1" 
```{r}
# Code independent variables to be "0" and "1" 
# Recoding the '2's to '0's is valid because the NSDUH codes these variables to be '2' if 'No' and '1' if 'Yes'.
# The following code will then code 'Yes' as '1' and 'No' as '0'
# Only Depression Variables are coded like this 
NSDUH2015to2019Data$amdeyr <- ifelse(NSDUH2015to2019Data$amdeyr == 2, 0, NSDUH2015to2019Data$amdeyr)
NSDUH2015to2019Data$amdelt <- ifelse(NSDUH2015to2019Data$amdelt == 2, 0, NSDUH2015to2019Data$amdelt)
NSDUH2015to2019Data$amdeimp <- ifelse(NSDUH2015to2019Data$amdeimp == 2, 0, NSDUH2015to2019Data$amdeimp)

```

```{r}
plyr::count(NSDUH2015to2019Data$spdmon)
plyr::count(NSDUH2015to2019Data$mhsuithk)
plyr::count(NSDUH2015to2019Data$mhsuipln)
plyr::count(NSDUH2015to2019Data$amdeyr)
plyr::count(NSDUH2015to2019Data$amdelt)
plyr::count(NSDUH2015to2019Data$amdeimp)
```


# Keep only the data with "heterosexual", "bisexual", or "gay/lesbian" responses
```{r}
NSDUH2015to2019Data <- NSDUH2015to2019Data %>%
  mutate(sexident = case_when(
    sexident == 1 ~ "Heterosexual",
    sexident == 2 ~ "Lesbian or Gay",
    sexident == 3 ~ "Bisexual",
    sexident == 85 ~ "BAD DATA Logically assigned",
    sexident == 94 ~ "DON'T KNOW",
    sexident == 97 ~ "REFUSED",
    sexident == 98 ~ "BLANK (NO ANSWER)",
    sexident == 99 ~ "LEGITIMATE SKIP",
    TRUE ~ as.character(sexident)  # Just in case there's like, other stuff
  )) %>%
  filter(sexident %in% c("Heterosexual", "Lesbian or Gay", "Bisexual"))  # Filter to keep only 1, 2, 3

plyr::count(NSDUH2015to2019Data$sexident) 
```
# Rename Covariates 
```{r}
NSDUH2015to2019Data <- NSDUH2015to2019Data %>%
  dplyr::rename(
    PREEducationVariable = IREDUHIGHST2,
    EcstasyVariable = ecstmoflag,
    InhalantVariable = inhalflag,
    PainRelieverVariable = pnranyflag,
    TranquilizerVariable = trqanyflag,
    StimulantVariable = stmanyflag,
    SedativeVariable = sedanyflag,
    KetamineVar = ketminflag,
    PoppersVariable = AMYLNIT2 # Ever INHALED AMYL NITRITE, "POPPERS", RUSH, ETC.
)
```


```{r}
####Counts of Variables (to confirm that there are no -9s or other null numerical values)####
plyr::count(NSDUH2015to2019Data$spdmon)
plyr::count(NSDUH2015to2019Data$mhsuithk)
plyr::count(NSDUH2015to2019Data$mhsuipln)
plyr::count(NSDUH2015to2019Data$amdeyr)
plyr::count(NSDUH2015to2019Data$amdelt)
plyr::count(NSDUH2015to2019Data$amdeimp)
plyr::count(NSDUH2015to2019Data$MarriageVariable)
plyr::count(NSDUH2015to2019Data$PREEducationVariable)
plyr::count(NSDUH2015to2019Data$EcstasyVariable)
plyr::count(NSDUH2015to2019Data$InhalantVariable)
plyr::count(NSDUH2015to2019Data$PainRelieverVariable)
plyr::count(NSDUH2015to2019Data$TranquilizerVariable)
plyr::count(NSDUH2015to2019Data$StimulantVariable)
plyr::count(NSDUH2015to2019Data$SedativeVariable)
plyr::count(NSDUH2015to2019Data$RiskyVariable)
plyr::count(NSDUH2015to2019Data$PSILCY2)
plyr::count(NSDUH2015to2019Data$PEYOTE2)
plyr::count(NSDUH2015to2019Data$lsdflag)
plyr::count(NSDUH2015to2019Data$MESC2)
plyr::count(NSDUH2015to2019Data$CATAG6)
plyr::count(NSDUH2015to2019Data$irsex)
plyr::count(NSDUH2015to2019Data$NEWRACE2)
plyr::count(NSDUH2015to2019Data$income)
plyr::count(NSDUH2015to2019Data$cocflag)
plyr::count(NSDUH2015to2019Data$herflag)
plyr::count(NSDUH2015to2019Data$pcpflag)
plyr::count(NSDUH2015to2019Data$mrjflag)
plyr::count(NSDUH2015to2019Data$KetamineVar)
plyr::count(NSDUH2015to2019Data$amdeimp)
plyr::count(NSDUH2015to2019Data$PoppersVariable)

 
```
# VARIABLE LABELING/CONVERSION INTO FACTORS 
```{r}
# Convert Lifetime Use Variables and Outcome Variables into Factors
# Overall - Main Dependent/Independent Variables
NSDUH2015to2019Data$spdmon <- as.factor(NSDUH2015to2019Data$spdmon)
NSDUH2015to2019Data$mhsuithk <- as.factor(NSDUH2015to2019Data$mhsuithk)
NSDUH2015to2019Data$mhsuipln <- as.factor(NSDUH2015to2019Data$mhsuipln)
NSDUH2015to2019Data$amdeyr <- as.factor(NSDUH2015to2019Data$amdeyr)
NSDUH2015to2019Data$amdelt <- as.factor(NSDUH2015to2019Data$amdelt)
NSDUH2015to2019Data$amdeimp <- as.factor(NSDUH2015to2019Data$amdeimp)
NSDUH2015to2019Data$EcstasyVariable <- as.factor(NSDUH2015to2019Data$EcstasyVariable)
NSDUH2015to2019Data$InhalantVariable <- as.factor(NSDUH2015to2019Data$InhalantVariable)
NSDUH2015to2019Data$PainRelieverVariable <- as.factor(NSDUH2015to2019Data$PainRelieverVariable)
NSDUH2015to2019Data$TranquilizerVariable <- as.factor(NSDUH2015to2019Data$TranquilizerVariable)
NSDUH2015to2019Data$StimulantVariable <- as.factor(NSDUH2015to2019Data$StimulantVariable)
NSDUH2015to2019Data$SedativeVariable <- as.factor(NSDUH2015to2019Data$SedativeVariable)
NSDUH2015to2019Data$PSILCY2 <- as.factor(NSDUH2015to2019Data$PSILCY2)
NSDUH2015to2019Data$PEYOTE2 <- as.factor(NSDUH2015to2019Data$PEYOTE2)
NSDUH2015to2019Data$lsdflag <- as.factor(NSDUH2015to2019Data$lsdflag)
NSDUH2015to2019Data$MESC2 <- as.factor(NSDUH2015to2019Data$MESC2)
NSDUH2015to2019Data$cocflag <- as.factor(NSDUH2015to2019Data$cocflag)
NSDUH2015to2019Data$herflag <- as.factor(NSDUH2015to2019Data$herflag)
NSDUH2015to2019Data$pcpflag <- as.factor(NSDUH2015to2019Data$pcpflag)
NSDUH2015to2019Data$mrjflag <- as.factor(NSDUH2015to2019Data$mrjflag)
NSDUH2015to2019Data$KetamineVar <- as.factor(NSDUH2015to2019Data$KetamineVar)
NSDUH2015to2019Data$PoppersVariable <- as.factor(NSDUH2015to2019Data$PoppersVariable)
NSDUH2015to2019Data$sexident <- as.factor(NSDUH2015to2019Data$sexident)
# Relevel SexIdent Variable 
NSDUH2015to2019Data$sexident <- relevel(NSDUH2015to2019Data$sexident, ref = "Heterosexual")
```
###Converting Demographic Factors/Covariates into Factors#####

```{r}
#Year
#Overall - Year
NSDUH2015to2019Data$year <- as.factor(NSDUH2015to2019Data$year)


###Marriage (Change Factors for All Four SO Data Sets)
#Overall - Marriage
NSDUH2015to2019Data$MarriageVariable <- factor(NSDUH2015to2019Data$MarriageVariable,
                                                  levels = c(1,2,3,4),
                                                  labels = c("Married", "Widowed", "Divorced or Separated", "Never Been Married"))

####Education
NSDUH2015to2019Data$EducationVariable <- ifelse(NSDUH2015to2019Data$PREEducationVariable<=4,1,
                                                   ifelse(NSDUH2015to2019Data$PREEducationVariable<=8, 2, 
                                                          ifelse(NSDUH2015to2019Data$PREEducationVariable >=9, 3,NSDUH2015to2019Data$PREEducationVariable)))

plyr::count(NSDUH2015to2019Data$EducationVariable)

NSDUH2015to2019Data$EducationVariable <- factor(NSDUH2015to2019Data$EducationVariable,
                                              levels = c(1,2,3),
                                              labels = c("Less than HS", "Some HS/HS Grad", "Some College or Above"))
plyr::count(NSDUH2015to2019Data$EducationVariable)

#Age
#Overall - Age
NSDUH2015to2019Data$CATAG6 <- factor(NSDUH2015to2019Data$CATAG6,
                                        levels = c(2,3,4,5),
                                        labels = c("18-25", "26-34", "35-49", "50+"))


###Sex
#Overall - Sex
NSDUH2015to2019Data$irsex <- factor(NSDUH2015to2019Data$irsex,
                                       levels = c(1,2),
                                       labels = c("Male", "Female"))

###Yearly Household Income
#Overall - Income
NSDUH2015to2019Data$income <- factor(NSDUH2015to2019Data$income,
                                        levels = c(1,2,3,4),
                                        labels = c("<$20,000", "$20,000-$49,999", "$50,000-$74,999","$75,000+"))
#Overall - Race
NSDUH2015to2019Data$NEWRACE2 <- factor(NSDUH2015to2019Data$NEWRACE2,
                                        levels = c(1,2,3,4,5,6,7),
                                        labels = c("NonHisp White", "NonHisp Black/Afr Am", "NonHisp Native Am/AK Native","NonHisp Native HI/Other Pac Isl","NonHisp Asian", "NonHisp more than one race", "Hispanic"))
#Self-Reported Engagement in Risky Behavior
#Overall
NSDUH2015to2019Data$RiskyVariable <- factor(NSDUH2015to2019Data$RiskyVariable,
                                           levels = c(1,2,3,4),
                                           labels = c("Never", "Seldom", "Sometimes", "Always"))

```

```{r}
plyr::count(NSDUH2015to2019Data$sexident)
```

## Split into subgroups, relabel, factor, and check counts
```{r}
# Split
# Pooled 5-year analysis weight (SAMHSA guidance for multi-year NSDUH): divide the single-year
# weight by the number of years. This is a CONSTANT rescaling, so it does not change the odds
# ratios, design-based standard errors, or weighted proportions; it only corrects absolute
# population-count estimates (not reported here).
NSDUH2015to2019Data$ANALWT_pooled <- NSDUH2015to2019Data$ANALWT_C / 5

NSDUH2015to2019DataHet <- subset(NSDUH2015to2019Data, sexident == 'Heterosexual')
NSDUH2015to2019DataBis <- subset(NSDUH2015to2019Data, sexident == 'Bisexual')
NSDUH2015to2019GayLesbian <- subset(NSDUH2015to2019Data, sexident == 'Lesbian or Gay')
```
```{r}

```


```{r}
#Relabel (Have to do this after the split, as the labels do not transfer over)
var_label(NSDUH2015to2019Data) <- list(PSILCY2 = "Psilocybin", EcstasyVariable = "MDMA/Ecstasy",
                                          lsdflag = "LSD", PEYOTE2 = "Peyote", MESC2 = "Mescaline",
                                          cocflag = "Cocaine", herflag = "Heroin",
                                          pcpflag = "PCP", InhalantVariable = "Inhalants",
                                          PainRelieverVariable = "Pain Relievers", TranquilizerVariable = "Tranquilizers",
                                          StimulantVariable = "Stimulants", SedativeVariable = "Sedatives",
                                          mrjflag = "Marijuana", NEWRACE2 = "Race/Ethnicity", spdmon = "Past Month Distress",  mhsuithk = "Past Yr. Suicidal Ideation",
                                          mhsuipln = "Past Yr. Suicidal Planning",  amdeyr = "Past Year MDE",amdelt = "Lifetime MDE", amdeimp = "Severe Impairment - MDE",
                                          MarriageVariable = "Marital Status", income = "Yearly Household Income",
                                          EducationVariable = "Educational Attainment", KetamineVar = "Ketamine", PoppersVariable = "Poppers") 

var_label(NSDUH2015to2019DataHet) <- list(PSILCY2 = "Psilocybin", EcstasyVariable = "MDMA/Ecstasy",
                                               lsdflag = "LSD", PEYOTE2 = "Peyote", MESC2 = "Mescaline",
                                               cocflag = "Cocaine", herflag = "Heroin",
                                               pcpflag = "PCP", InhalantVariable = "Inhalants",
                                               PainRelieverVariable = "Pain Relievers", TranquilizerVariable = "Tranquilizers",
                                               StimulantVariable = "Stimulants", SedativeVariable = "Sedatives",
                                               mrjflag = "Marijuana", NEWRACE2 = "Race/Ethnicity", spdmon = "Past Month Distress",  mhsuithk = "Past Yr. Suicidal Ideation",
                                               mhsuipln = "Past Yr. Suicidal Planning", amdeyr = "Past Year MDE",amdelt = "Lifetime MDE", amdeimp = "Severe Impairment - MDE",
                                               MarriageVariable = "Marital Status", income = "Yearly Household Income",
                                               EducationVariable = "Educational Attainment",KetamineVar = "Ketamine", PoppersVariable = "Poppers") 

var_label(NSDUH2015to2019DataBis) <- list(PSILCY2 = "Psilocybin", EcstasyVariable = "MDMA/Ecstasy",
                                               lsdflag = "LSD", PEYOTE2 = "Peyote", MESC2 = "Mescaline",
                                               cocflag = "Cocaine", herflag = "Heroin",
                                               pcpflag = "PCP", InhalantVariable = "Inhalants",
                                               PainRelieverVariable = "Pain Relievers", TranquilizerVariable = "Tranquilizers",
                                               StimulantVariable = "Stimulants", SedativeVariable = "Sedatives",
                                               mrjflag = "Marijuana", NEWRACE2 = "Race/Ethnicity", spdmon = "Past Month Distress",  mhsuithk = "Past Yr. Suicidal Ideation",
                                               mhsuipln = "Past Yr. Suicidal Planning", amdeyr = "Past Year MDE",amdelt = "Lifetime MDE", amdeimp = "Severe Impairment - MDE",
                                               MarriageVariable = "Marital Status", income = "Yearly Household Income",
                                               EducationVariable = "Educational Attainment",KetamineVar = "Ketamine", PoppersVariable = "Poppers") 

var_label(NSDUH2015to2019GayLesbian) <- list(PSILCY2 = "Psilocybin", EcstasyVariable = "MDMA/Ecstasy",
                                                                lsdflag = "LSD", PEYOTE2 = "Peyote", MESC2 = "Mescaline",
                                                                cocflag = "Cocaine", herflag = "Heroin",
                                                                pcpflag = "PCP", InhalantVariable = "Inhalants",
                                                                PainRelieverVariable = "Pain Relievers", TranquilizerVariable = "Tranquilizers",
                                                                StimulantVariable = "Stimulants", SedativeVariable = "Sedatives",
                                                                mrjflag = "Marijuana", NEWRACE2 = "Race/Ethnicity", spdmon = "Past Month Distress",  mhsuithk = "Past Yr. Suicidal Ideation",
                                                                mhsuipln = "Past Yr. Suicidal Planning", amdeyr = "Past Year MDE",amdelt = "Lifetime MDE", amdeimp = "Severe Impairment - MDE",
                                                                MarriageVariable = "Marital Status", income = "Yearly Household Income",
                                                                EducationVariable = "Educational Attainment", KetamineVar = "Ketamine", PoppersVariable = "Poppers") 

```
```{r}
plyr::count(NSDUH2015to2019Data$spdmon)
plyr::count(NSDUH2015to2019Data$mhsuithk)
plyr::count(NSDUH2015to2019Data$mhsuipln)
plyr::count(NSDUH2015to2019Data$amdeyr)
plyr::count(NSDUH2015to2019Data$amdelt)
plyr::count(NSDUH2015to2019Data$amdeimp)
```

Checking Variables for '-9' responses

```{r}
#### Re-Counts of Variables (to confirm that there are no -9s or other null numerical values) ####
plyr::count(NSDUH2015to2019Data$spdmon)
plyr::count(NSDUH2015to2019Data$mhsuithk)
plyr::count(NSDUH2015to2019Data$mhsuipln)
plyr::count(NSDUH2015to2019Data$amdeyr)
plyr::count(NSDUH2015to2019Data$amdelt)
plyr::count(NSDUH2015to2019Data$amdeimp)
plyr::count(NSDUH2015to2019Data$MarriageVariable)
plyr::count(NSDUH2015to2019Data$PREEducationVariable)
plyr::count(NSDUH2015to2019Data$EcstasyVariable)
plyr::count(NSDUH2015to2019Data$InhalantVariable)
plyr::count(NSDUH2015to2019Data$PainRelieverVariable)
plyr::count(NSDUH2015to2019Data$TranquilizerVariable)
plyr::count(NSDUH2015to2019Data$StimulantVariable)
plyr::count(NSDUH2015to2019Data$SedativeVariable)
plyr::count(NSDUH2015to2019Data$RiskyVariable)
plyr::count(NSDUH2015to2019Data$PSILCY2)
plyr::count(NSDUH2015to2019Data$PEYOTE2)
plyr::count(NSDUH2015to2019Data$lsdflag)
plyr::count(NSDUH2015to2019Data$MESC2)
plyr::count(NSDUH2015to2019Data$CATAG6)
plyr::count(NSDUH2015to2019Data$irsex)
plyr::count(NSDUH2015to2019Data$NEWRACE2)
plyr::count(NSDUH2015to2019Data$income)
plyr::count(NSDUH2015to2019Data$cocflag)
plyr::count(NSDUH2015to2019Data$herflag)
plyr::count(NSDUH2015to2019Data$pcpflag)
plyr::count(NSDUH2015to2019Data$mrjflag)
plyr::count(NSDUH2015to2019Data$KetamineVar)
plyr::count(NSDUH2015to2019Data$amdeimp)
plyr::count(NSDUH2015to2019Data$PoppersVariable)
```

```{r}
#### Counts of Variables (to confirm that there are no -9s or other null numerical values) ####
plyr::count(NSDUH2015to2019DataBis$spdmon)
plyr::count(NSDUH2015to2019DataBis$mhsuithk)
plyr::count(NSDUH2015to2019DataBis$mhsuipln)
plyr::count(NSDUH2015to2019DataBis$amdeyr)
plyr::count(NSDUH2015to2019DataBis$amdelt)
plyr::count(NSDUH2015to2019DataBis$amdeimp)
plyr::count(NSDUH2015to2019DataBis$MarriageVariable)
plyr::count(NSDUH2015to2019DataBis$PREEducationVariable)
plyr::count(NSDUH2015to2019DataBis$EcstasyVariable)
plyr::count(NSDUH2015to2019DataBis$InhalantVariable)
plyr::count(NSDUH2015to2019DataBis$PainRelieverVariable)
plyr::count(NSDUH2015to2019DataBis$TranquilizerVariable)
plyr::count(NSDUH2015to2019DataBis$StimulantVariable)
plyr::count(NSDUH2015to2019DataBis$SedativeVariable)
plyr::count(NSDUH2015to2019DataBis$RiskyVariable)
plyr::count(NSDUH2015to2019DataBis$PSILCY2)
plyr::count(NSDUH2015to2019DataBis$PEYOTE2)
plyr::count(NSDUH2015to2019DataBis$lsdflag)
plyr::count(NSDUH2015to2019DataBis$MESC2)
plyr::count(NSDUH2015to2019DataBis$CATAG6)
plyr::count(NSDUH2015to2019DataBis$irsex)
plyr::count(NSDUH2015to2019DataBis$NEWRACE2)
plyr::count(NSDUH2015to2019DataBis$income)
plyr::count(NSDUH2015to2019DataBis$cocflag)
plyr::count(NSDUH2015to2019DataBis$herflag)
plyr::count(NSDUH2015to2019DataBis$pcpflag)
plyr::count(NSDUH2015to2019DataBis$mrjflag)
plyr::count(NSDUH2015to2019DataBis$KetamineVar)
plyr::count(NSDUH2015to2019DataBis$amdeimp)
plyr::count(NSDUH2015to2019DataBis$PoppersVariable)

```

```{r}
#### Counts of Variables (to confirm that there are no -9s or other null numerical values) ####
plyr::count(NSDUH2015to2019GayLesbian$spdmon)
plyr::count(NSDUH2015to2019GayLesbian$mhsuithk)
plyr::count(NSDUH2015to2019GayLesbian$mhsuipln)
plyr::count(NSDUH2015to2019GayLesbian$amdeyr)
plyr::count(NSDUH2015to2019GayLesbian$amdelt)
plyr::count(NSDUH2015to2019GayLesbian$amdeimp)
plyr::count(NSDUH2015to2019GayLesbian$MarriageVariable)
plyr::count(NSDUH2015to2019GayLesbian$PREEducationVariable)
plyr::count(NSDUH2015to2019GayLesbian$EcstasyVariable)
plyr::count(NSDUH2015to2019GayLesbian$InhalantVariable)
plyr::count(NSDUH2015to2019GayLesbian$PainRelieverVariable)
plyr::count(NSDUH2015to2019GayLesbian$TranquilizerVariable)
plyr::count(NSDUH2015to2019GayLesbian$StimulantVariable)
plyr::count(NSDUH2015to2019GayLesbian$SedativeVariable)
plyr::count(NSDUH2015to2019GayLesbian$RiskyVariable)
plyr::count(NSDUH2015to2019GayLesbian$PSILCY2)
plyr::count(NSDUH2015to2019GayLesbian$PEYOTE2)
plyr::count(NSDUH2015to2019GayLesbian$lsdflag)
plyr::count(NSDUH2015to2019GayLesbian$MESC2)
plyr::count(NSDUH2015to2019GayLesbian$CATAG6)
plyr::count(NSDUH2015to2019GayLesbian$irsex)
plyr::count(NSDUH2015to2019GayLesbian$NEWRACE2)
plyr::count(NSDUH2015to2019GayLesbian$income)
plyr::count(NSDUH2015to2019GayLesbian$cocflag)
plyr::count(NSDUH2015to2019GayLesbian$herflag)
plyr::count(NSDUH2015to2019GayLesbian$pcpflag)
plyr::count(NSDUH2015to2019GayLesbian$mrjflag)
plyr::count(NSDUH2015to2019GayLesbian$KetamineVar)
plyr::count(NSDUH2015to2019GayLesbian$amdeimp)
plyr::count(NSDUH2015to2019GayLesbian$PoppersVariable)

```
```{r}
#### Counts of Variables (to confirm that there are no -9s or other null numerical values) ####
plyr::count(NSDUH2015to2019DataHet$spdmon)
plyr::count(NSDUH2015to2019DataHet$mhsuithk)
plyr::count(NSDUH2015to2019DataHet$mhsuipln)
plyr::count(NSDUH2015to2019DataHet$amdeyr)
plyr::count(NSDUH2015to2019DataHet$amdelt)
plyr::count(NSDUH2015to2019DataHet$amdeimp)
plyr::count(NSDUH2015to2019DataHet$MarriageVariable)
plyr::count(NSDUH2015to2019DataHet$PREEducationVariable)
plyr::count(NSDUH2015to2019DataHet$EcstasyVariable)
plyr::count(NSDUH2015to2019DataHet$InhalantVariable)
plyr::count(NSDUH2015to2019DataHet$PainRelieverVariable)
plyr::count(NSDUH2015to2019DataHet$TranquilizerVariable)
plyr::count(NSDUH2015to2019DataHet$StimulantVariable)
plyr::count(NSDUH2015to2019DataHet$SedativeVariable)
plyr::count(NSDUH2015to2019DataHet$RiskyVariable)
plyr::count(NSDUH2015to2019DataHet$PSILCY2)
plyr::count(NSDUH2015to2019DataHet$PEYOTE2)
plyr::count(NSDUH2015to2019DataHet$lsdflag)
plyr::count(NSDUH2015to2019DataHet$MESC2)
plyr::count(NSDUH2015to2019DataHet$CATAG6)
plyr::count(NSDUH2015to2019DataHet$irsex)
plyr::count(NSDUH2015to2019DataHet$NEWRACE2)
plyr::count(NSDUH2015to2019DataHet$income)
plyr::count(NSDUH2015to2019DataHet$cocflag)
plyr::count(NSDUH2015to2019DataHet$herflag)
plyr::count(NSDUH2015to2019DataHet$pcpflag)
plyr::count(NSDUH2015to2019DataHet$mrjflag)
plyr::count(NSDUH2015to2019DataHet$KetamineVar)
plyr::count(NSDUH2015to2019DataHet$amdeimp)
plyr::count(NSDUH2015to2019DataHet$PoppersVariable)

```


```{r}
####NSDUH SURVEY DESIGN###
 #Now, it is time to incorporate the survey weighting. I use the survey package to specify weights
 #vestr = strata; verep = cluster/id variable
 #This link details how to include survey data: https://stylizeddata.com/how-to-use-survey-weights-in-r/
 #Overall
totalDesign <- survey::svydesign(id      = ~verep,
                                      strata  = ~vestr,
                                      weights = ~ANALWT_pooled,
                                      nest    = TRUE,
                                      data    = NSDUH2015to2019Data)


hetDesign <- survey::svydesign(id      = ~verep,
                                      strata  = ~vestr,
                                      weights = ~ANALWT_pooled,
                                      nest    = TRUE,
                                      data    = NSDUH2015to2019DataHet)


biDesign <- survey::svydesign(id      = ~verep,
                                               strata  = ~vestr,
                                               weights = ~ANALWT_pooled,
                                               nest    = TRUE,
                                               data    = NSDUH2015to2019DataBis)


lezDesign <- survey::svydesign(id      = ~verep,
                                           strata  = ~vestr,
                                           weights = ~ANALWT_pooled,
                                           nest    = TRUE,
                                           data    = NSDUH2015to2019GayLesbian)


```


#######################SECTION 1: OVERALL SAMPLE ECSTASY/PSYCHEDELIC USE ANALYSES###################
```{r}
# Exploring the data - ensuring sufficient sample size 
 #  Count occurrences of psilocybin use and calculate percentages within each sexident group
NSDUH2015to2019Data %>%
  group_by(sexident, PSILCY2) %>%
  dplyr::summarise(count = n(), .groups = 'drop') %>%
  # Calculate the percentage within each sexident group
  group_by(sexident) %>%
  mutate(percentage = (count / sum(count)) * 100) %>%
  arrange(sexident, PSILCY2)

 # Count occurrences of Ecstasy Use by sexident
 # Count occurrences and calculate percentages within each sexident group
NSDUH2015to2019Data %>%
  group_by(sexident, EcstasyVariable) %>%
  dplyr::summarise(count = n(), .groups = 'drop') %>%
  # Calculate the percentage within each sexident group
  group_by(sexident) %>%
  mutate(percentage = (count / sum(count)) * 100) %>%
  arrange(sexident, EcstasyVariable)
```
PSYCHOLOGICAL DISTRESS

### PSY DIS Overall
```{r} 

PSYDISTMODEL2015to2019 <- survey::svyglm(spdmon ~ EcstasyVariable + PSILCY2+ lsdflag + PEYOTE2 +MESC2 + MarriageVariable+  EducationVariable+ CATAG6+irsex+ income+ NEWRACE2+ sexident +KetamineVar +  PoppersVariable+
                                 RiskyVariable+ cocflag+ herflag+ pcpflag+  InhalantVariable + PainRelieverVariable+ TranquilizerVariable +
                                 StimulantVariable+ SedativeVariable+ mrjflag+year, data=NSDUH2015to2019Data,family=quasibinomial(), design=totalDesign)
summary(PSYDISTMODEL2015to2019)
DistressModelOdds2015to2019 <- exp(cbind(OR = coef(PSYDISTMODEL2015to2019), or_confint(PSYDISTMODEL2015to2019)))
DistressModelOdds2015to2019
DistressModelOdds2015to2019 <- round(DistressModelOdds2015to2019, digits =2)
DistressModelOdds2015to2019[c(2,3),]
```

### Psilocybin SO Interaction
```{r}

PSYDISTMODEL2015to2019INTERpsilthree <- survey::svyglm(spdmon ~ EcstasyVariable + sexident*PSILCY2 + cocflag+  lsdflag  + KetamineVar + PoppersVariable+ PEYOTE2 +MESC2 + MarriageVariable+ EducationVariable+ CATAG6+ NEWRACE2 +irsex+ income+
                                                RiskyVariable+ herflag+ pcpflag+  InhalantVariable + PainRelieverVariable+ TranquilizerVariable+
                                                StimulantVariable+ SedativeVariable+ mrjflag+year, data=NSDUH2015to2019Data,family=quasibinomial(), design=totalDesign)
summary(PSYDISTMODEL2015to2019INTERpsilthree)
```


# Suicidal Ideation 
### Suicidal Ideation Overall 
```{r}
# Suicidal Ideation 
### Suicidal Ideation Overall 
SUITHINKMODEL2015to2019 <- survey::svyglm(mhsuithk ~ EcstasyVariable + PSILCY2+ lsdflag + KetamineVar +  PoppersVariable+ PEYOTE2 +MESC2 + MarriageVariable+EducationVariable+ CATAG6+  sexident +irsex+ NEWRACE2+ income+
                                 RiskyVariable+ cocflag+ herflag+ pcpflag+  InhalantVariable + PainRelieverVariable+ TranquilizerVariable+
                                 StimulantVariable+ SedativeVariable+ mrjflag+year, data=NSDUH2015to2019Data,family=quasibinomial(), design=totalDesign)
              
summary(SUITHINKMODEL2015to2019)
SUITHINKModelOdds2015to2019 <- exp(cbind(OR = coef(SUITHINKMODEL2015to2019), or_confint(SUITHINKMODEL2015to2019)))
SUITHINKModelOdds2015to2019
SUITHINKModelOdds2015to2019 <- round(SUITHINKModelOdds2015to2019, digits =2)
SUITHINKModelOdds2015to2019[c(2:4),]
             
```
### Psilocybin SO Interaction
```{r}

SUITHINKMODEL2015to2019INTERpsilthree <- survey::svyglm(mhsuithk ~ EcstasyVariable+ sexident*PSILCY2 + lsdflag + PEYOTE2 +MESC2 + MarriageVariable+ EducationVariable+ CATAG6+  KetamineVar + PoppersVariable + irsex+ income+NEWRACE2 + 
                                                     RiskyVariable+ cocflag+ herflag+ pcpflag+  InhalantVariable + PainRelieverVariable+ TranquilizerVariable+
                                                     StimulantVariable+ SedativeVariable+ mrjflag+year, data=NSDUH2015to2019Data,family=quasibinomial(), design=totalDesign)

summary(SUITHINKMODEL2015to2019INTERpsilthree)
SUITHINKMODEL2015to2019INTERpsilthree_OR <- or_table(SUITHINKMODEL2015to2019INTERpsilthree)
SUITHINKMODEL2015to2019INTERpsilthree_OR

```
### MDMA SO Interaction
```{r}

SUITHINKMODEL2015to2019INTERecstthree <- survey::svyglm(mhsuithk ~ sexident*EcstasyVariable+ PSILCY2 + KetamineVar +  PoppersVariable+ lsdflag + PEYOTE2 +MESC2 + MarriageVariable+ EducationVariable+CATAG6 +  irsex+ income+NEWRACE2 +
                                                     RiskyVariable+ cocflag+ herflag+ pcpflag+  InhalantVariable + PainRelieverVariable+ TranquilizerVariable+
                                                     StimulantVariable+ SedativeVariable+ mrjflag+year, data=NSDUH2015to2019Data,family=quasibinomial(), design=totalDesign)

summary(SUITHINKMODEL2015to2019INTERecstthree)
SUITHINKMODEL2015to2019INTERecstthree_OR <- or_table(SUITHINKMODEL2015to2019INTERecstthree)
SUITHINKMODEL2015to2019INTERecstthree_OR
```
## Significant interaction between bisexual identity x psilocybin use
## So, I proceed with subgroup analysis
## Suicidal Ideation by Sexual Identity 
### Heterosexual
```{r}
## Suicidal Ideation by Sexual Identity 
### Heterosexual
SUITHINKMODEL2015to2019HET <- survey::svyglm(mhsuithk ~ EcstasyVariable + PSILCY2+ lsdflag + KetamineVar +  PoppersVariable+ PEYOTE2 +MESC2 + MarriageVariable+ EducationVariable+ CATAG6+ NEWRACE2+ irsex+  income+
                                           RiskyVariable+ cocflag+ herflag+ pcpflag+  InhalantVariable + PainRelieverVariable+ TranquilizerVariable+
                                           StimulantVariable+ SedativeVariable+ mrjflag+year, data=NSDUH2015to2019DataHet,family=quasibinomial(), design=hetDesign)
summary(SUITHINKMODEL2015to2019HET)
SUITHINKModelOdds2015to2019HET <- exp(cbind(OR = coef(SUITHINKMODEL2015to2019HET), or_confint(SUITHINKMODEL2015to2019HET)))
SUITHINKModelOdds2015to2019HET
SUITHINKModelOdds2015to2019HET <- round(SUITHINKModelOdds2015to2019HET, digits =2)
SUITHINKModelOdds2015to2019HET[c(2,3),]
```
# Gay and Lesbian
```{r}

SUITHINKMODEL2015to2019Lez <- survey::svyglm(mhsuithk ~ EcstasyVariable + PSILCY2+ lsdflag + PEYOTE2 +MESC2 + KetamineVar +  PoppersVariable+MarriageVariable+ EducationVariable+ CATAG6+ NEWRACE2+ irsex+  income+
                                           RiskyVariable+ cocflag+ herflag+ pcpflag+  InhalantVariable + PainRelieverVariable+ TranquilizerVariable+
                                           StimulantVariable+ SedativeVariable+ mrjflag+year, data=NSDUH2015to2019GayLesbian,family=quasibinomial(), design=lezDesign)
summary(SUITHINKMODEL2015to2019Lez)
SUITHINKModelOdds2015to2019Lez <- exp(cbind(OR = coef(SUITHINKMODEL2015to2019Lez), or_confint(SUITHINKMODEL2015to2019Lez)))
SUITHINKModelOdds2015to2019Lez
SUITHINKModelOdds2015to2019Lez <- round(SUITHINKModelOdds2015to2019Lez, digits =2)
SUITHINKModelOdds2015to2019Lez[c(2,3),]

```
# Bisexual
```{r}


SUITHINKMODEL2015to2019Bi <- survey::svyglm(mhsuithk ~ EcstasyVariable + PSILCY2+ lsdflag + PEYOTE2 +MESC2 + MarriageVariable+ EducationVariable+ CATAG6+ NEWRACE2+ irsex+  income+
                                           RiskyVariable+ cocflag+ herflag+ KetamineVar +  PoppersVariable+pcpflag+  InhalantVariable + PainRelieverVariable+ TranquilizerVariable+
                                           StimulantVariable+ SedativeVariable+ mrjflag+year, data=NSDUH2015to2019DataBis,family=quasibinomial(), design=biDesign)
summary(SUITHINKMODEL2015to2019Bi)
SUITHINKModelOdds2015to2019Bi <- exp(cbind(OR = coef(SUITHINKMODEL2015to2019Bi), or_confint(SUITHINKMODEL2015to2019Bi)))
SUITHINKModelOdds2015to2019Bi
SUITHINKModelOdds2015to2019Bi <- round(SUITHINKModelOdds2015to2019Bi, digits =2)
SUITHINKModelOdds2015to2019Bi[c(2,3),]
```


# Suicidal Planning
### Suicidal Planning Overall 
```{r}
# Suicidal Planning
### Suicidal Planning Overall 

SUIPLANMODEL2015to2019 <- survey::svyglm(mhsuipln ~ EcstasyVariable + PSILCY2+ lsdflag + PEYOTE2 +MESC2 + MarriageVariable+ EducationVariable+ CATAG6+ sexident + irsex+ NEWRACE2+ income+ KetamineVar +  PoppersVariable+ 
                                 RiskyVariable+ cocflag+ herflag+ pcpflag+  InhalantVariable + PainRelieverVariable+ TranquilizerVariable+
                                 StimulantVariable+ SedativeVariable+ mrjflag+year, data=NSDUH2015to2019Data,family=quasibinomial(), design=totalDesign)
summary(SUIPLANMODEL2015to2019)
SUIPLANModelOdds2015to2019 <- exp(cbind(OR = coef(SUIPLANMODEL2015to2019), or_confint(SUIPLANMODEL2015to2019)))
SUIPLANModelOdds2015to2019
SUIPLANModelOdds2015to2019 <- round(SUIPLANModelOdds2015to2019, digits =2)
SUIPLANModelOdds2015to2019[c(2:4),]
```

### MDMA SO Interaction
```{r}

SUICPLANMODEL2015to2019INTERecstthree <- survey::svyglm(mhsuipln ~ sexident*EcstasyVariable+cocflag+ PSILCY2 + lsdflag + PEYOTE2 +MESC2 + MarriageVariable+ EducationVariable+CATAG6+ NEWRACE2 +  irsex+ income+ KetamineVar + PoppersVariable+ 
                                                     RiskyVariable+ herflag+ pcpflag+  InhalantVariable + PainRelieverVariable+ TranquilizerVariable+
                                                     StimulantVariable+ SedativeVariable+ mrjflag+year, data=NSDUH2015to2019Data,family=quasibinomial(), design=totalDesign)

summary(SUICPLANMODEL2015to2019INTERecstthree)
SUICPLANMODEL2015to2019INTERecstthree_OR <- or_table(SUICPLANMODEL2015to2019INTERecstthree)
SUICPLANMODEL2015to2019INTERecstthree_OR
```


## Past year MDE
# Past year MDE
```{r}
# Past Year MDE
###  Past Year MDE Overall 

pstyrMDEMODEL2015to2019 <- survey::svyglm(amdeyr ~ EcstasyVariable + PSILCY2+ lsdflag + PEYOTE2 +MESC2 + MarriageVariable+ EducationVariable+ CATAG6+ irsex + sexident + PoppersVariable+KetamineVar +NEWRACE2+ income+  
                                 RiskyVariable+ cocflag+ herflag+ pcpflag+  InhalantVariable + PainRelieverVariable+ TranquilizerVariable+
                                 StimulantVariable+ SedativeVariable+ mrjflag+year, data=NSDUH2015to2019Data,family=quasibinomial(), design=totalDesign)
```

```{r}
summary(pstyrMDEMODEL2015to2019)
pstyrMDEModelOdds2015to2019 <- exp(cbind(OR = coef(pstyrMDEMODEL2015to2019), or_confint(pstyrMDEMODEL2015to2019)))
pstyrMDEModelOdds2015to2019
pstyrMDEModelOdds2015to2019 <- round(pstyrMDEModelOdds2015to2019, digits =2)
pstyrMDEModelOdds2015to2019[c(2:4),]
```
### Psilocybin SO Interaction
```{r}


pstyrMDEMODEL2015to2019INTERpsilthree <- survey::svyglm(amdeyr ~ EcstasyVariable+ sexident*PSILCY2 +cocflag+ lsdflag + PEYOTE2 +MESC2 + MarriageVariable+ EducationVariable+ CATAG6+ NEWRACE2 + irsex+ income+
                                                     RiskyVariable+ herflag+ pcpflag+  InhalantVariable + PainRelieverVariable+ TranquilizerVariable+KetamineVar +  PoppersVariable+ 
                                                     StimulantVariable+ SedativeVariable+ mrjflag+year, data=NSDUH2015to2019Data,family=quasibinomial(), design=totalDesign)

summary(pstyrMDEMODEL2015to2019INTERpsilthree)
pstyrMDEMODEL2015to2019INTERpsilthree_OR <- or_table(pstyrMDEMODEL2015to2019INTERpsilthree)
pstyrMDEMODEL2015to2019INTERpsilthree_OR
```
### MDMA SO Interaction
```{r}

pstyrMDEMODEL2015to2019INTERecstthree <- survey::svyglm(amdeyr ~ sexident*EcstasyVariable+ PSILCY2 + lsdflag + PEYOTE2 +MESC2 + MarriageVariable+ EducationVariable+CATAG6+ NEWRACE2 +  irsex+ income+
                                                     RiskyVariable+ cocflag+ herflag+ pcpflag+  InhalantVariable + PainRelieverVariable+ TranquilizerVariable+KetamineVar +  PoppersVariable+ 
                                                     StimulantVariable+ SedativeVariable+ mrjflag+year, data=NSDUH2015to2019Data,family=quasibinomial(), design=totalDesign)

summary(pstyrMDEMODEL2015to2019INTERecstthree)
pstyrMDEMODEL2015to2019INTERecstthree_OR <- or_table(pstyrMDEMODEL2015to2019INTERecstthree)
pstyrMDEMODEL2015to2019INTERecstthree_OR
```

# Lifetime MDE
### Lifetime MDE Overall 
```{r}


lifetimeMDEMODEL2015to2019 <- survey::svyglm(amdelt ~ EcstasyVariable + PSILCY2+ lsdflag + PEYOTE2 +MESC2 + MarriageVariable+ EducationVariable+ CATAG6+ sexident + irsex+ NEWRACE2+ income+
                                 RiskyVariable+ cocflag+ herflag+ pcpflag+ InhalantVariable + PainRelieverVariable+ TranquilizerVariable+KetamineVar +PoppersVariable+ 
                                 StimulantVariable+ SedativeVariable+ mrjflag+year, data=NSDUH2015to2019Data,family=quasibinomial(), design=totalDesign)

```
```{r}
summary(lifetimeMDEMODEL2015to2019)
lifetimeMDEModelOdds2015to2019 <- exp(cbind(OR = coef(lifetimeMDEMODEL2015to2019), or_confint(lifetimeMDEMODEL2015to2019)))
lifetimeMDEModelOdds2015to2019
lifetimeMDEModelOdds2015to2019 <- round(lifetimeMDEModelOdds2015to2019, digits = 2)
lifetimeMDEModelOdds2015to2019[c(2:4),]
```
### Psilocybin SO Interaction
```{r}

lifetimeMDEMODEL2015to2019INTERpsilthree <- survey::svyglm(amdelt ~ EcstasyVariable+ sexident*PSILCY2 + lsdflag + PEYOTE2 +MESC2 + cocflag+MarriageVariable+ EducationVariable+ CATAG6+ NEWRACE2 + irsex+ income+
                                                     RiskyVariable+  herflag+ pcpflag+ InhalantVariable + PainRelieverVariable+ TranquilizerVariable+KetamineVar +  PoppersVariable+ 
                                                     StimulantVariable+ SedativeVariable+ mrjflag+year, data=NSDUH2015to2019Data,family=quasibinomial(), design=totalDesign)

summary(lifetimeMDEMODEL2015to2019INTERpsilthree)
lifetimeMDEMODEL2015to2019INTERpsilthree_OR <- or_table(lifetimeMDEMODEL2015to2019INTERpsilthree)
lifetimeMDEMODEL2015to2019INTERpsilthree_OR
```
### MDMA SO Interaction
```{r}


lifetimeMDEMODEL2015to2019INTERecstthree <- survey::svyglm(amdelt ~ sexident*EcstasyVariable+ PSILCY2 + lsdflag + PEYOTE2 +MESC2 + MarriageVariable+ EducationVariable+CATAG6+ NEWRACE2 + irsex+ income+
                                                     RiskyVariable+ cocflag+ herflag+ pcpflag+ InhalantVariable + PainRelieverVariable+ TranquilizerVariable+KetamineVar + PoppersVariable+ 
                                                     StimulantVariable+ SedativeVariable+ mrjflag+year, data=NSDUH2015to2019Data,family=quasibinomial(), design=totalDesign)

summary(lifetimeMDEMODEL2015to2019INTERecstthree)
lifetimeMDEMODEL2015to2019INTERecstthree_OR <- or_table(lifetimeMDEMODEL2015to2019INTERecstthree)
lifetimeMDEMODEL2015to2019INTERecstthree_OR
```

# Severe MDE
### Severe MDE Overall
```{r}

severeMDEMODEL2015to2019 <- survey::svyglm(amdeimp ~ EcstasyVariable + PSILCY2+ lsdflag + PEYOTE2 +MESC2 + MarriageVariable+ EducationVariable+ CATAG6+ sexident + irsex+NEWRACE2+ income+
                                 RiskyVariable+ cocflag+ herflag+ pcpflag+ InhalantVariable + PainRelieverVariable+ TranquilizerVariable+KetamineVar +  PoppersVariable+ 
                                 StimulantVariable+ SedativeVariable+ mrjflag+year, data=NSDUH2015to2019Data,family=quasibinomial(), design=totalDesign)
summary(severeMDEMODEL2015to2019)
severeMDEModelOdds2015to2019 <- exp(cbind(OR = coef(severeMDEMODEL2015to2019), or_confint(severeMDEMODEL2015to2019)))
severeMDEModelOdds2015to2019
severeMDEModelOdds2015to2019 <- round(severeMDEModelOdds2015to2019, digits = 2)
severeMDEModelOdds2015to2019[c(2:4),]

```
### Psilocybin SO Interaction
```{r}

severeMDEMODEL2015to2019INTERpsilthree <- survey::svyglm(amdeimp ~ EcstasyVariable+ sexident* PSILCY2 + cocflag+lsdflag + PEYOTE2 +MESC2 + MarriageVariable+ EducationVariable+ CATAG6+ NEWRACE2 + irsex+ income+ RiskyVariable+ cocflag+ herflag+ pcpflag+ InhalantVariable + PainRelieverVariable+ TranquilizerVariable+KetamineVar +  PoppersVariable+ StimulantVariable+ SedativeVariable+ mrjflag+year, data=NSDUH2015to2019Data ,family=quasibinomial(), design=totalDesign)

summary(severeMDEMODEL2015to2019INTERpsilthree)
severeMDEMODEL2015to2019INTERpsilthree_OR <- or_table(severeMDEMODEL2015to2019INTERpsilthree)
severeMDEMODEL2015to2019INTERpsilthree_OR
```
### MDMA SO Interaction
```{r}

severeMDEMODEL2015to2019INTERecstthree <- survey::svyglm(amdeimp ~ sexident*EcstasyVariable+ PSILCY2 + lsdflag + PEYOTE2 +MESC2 + MarriageVariable+ EducationVariable+CATAG6+ NEWRACE2 + irsex+ income+RiskyVariable+ cocflag+ herflag+ pcpflag+ InhalantVariable + PainRelieverVariable+ TranquilizerVariable+KetamineVar + PoppersVariable+ 
StimulantVariable+ SedativeVariable+ mrjflag+year, data=NSDUH2015to2019Data,family=quasibinomial(), design= totalDesign)

summary(severeMDEMODEL2015to2019INTERecstthree)
severeMDEMODEL2015to2019INTERecstthree_OR <- or_table(severeMDEMODEL2015to2019INTERecstthree)
severeMDEMODEL2015to2019INTERecstthree_OR
```
## Significant interaction found between bisexual identity and lifetime psilocybin use
## So, we proceed with subgroup analysis
## Severe MDE by Sexual Identity 
### Heterosexual 
```{r}


severeMDEMODEL2015to2019HET <- survey::svyglm(amdeimp ~ EcstasyVariable + PSILCY2+ lsdflag + PEYOTE2 +MESC2 + MarriageVariable+ EducationVariable+ CATAG6+ NEWRACE2+ irsex+ income+
                                           RiskyVariable+ cocflag+ herflag+ pcpflag+ InhalantVariable + PainRelieverVariable+ TranquilizerVariable+KetamineVar +  PoppersVariable+ 
                                           StimulantVariable+ SedativeVariable+ mrjflag+year, data=NSDUH2015to2019DataHet,family=quasibinomial(), design=hetDesign)
summary(severeMDEMODEL2015to2019HET)
severeMDEMODEL2015to2019HET_OR <- or_table(severeMDEMODEL2015to2019HET)   # display only; keeps severeMDEMODEL2015to2019HET as the svyglm object
severeMDEMODEL2015to2019HET_OR
```
### Gay and Lesbian
```{r}

severeMDEMODEL2015to2019LG <- survey::svyglm(amdeimp ~ EcstasyVariable + PSILCY2+ lsdflag + PEYOTE2 +MESC2 + MarriageVariable+ EducationVariable+ CATAG6+ NEWRACE2+ irsex+ income+
                                           RiskyVariable+ cocflag+ herflag+ pcpflag+ InhalantVariable + PainRelieverVariable+ TranquilizerVariable+ KetamineVar +  PoppersVariable+ 
                                           StimulantVariable+ SedativeVariable+ mrjflag+year, data=NSDUH2015to2019GayLesbian,family=quasibinomial(), design=lezDesign)

summary(severeMDEMODEL2015to2019LG)
severeMDEMODEL2015to2019LG_OR <- or_table(severeMDEMODEL2015to2019LG)   # display only; keeps severeMDEMODEL2015to2019LG as the svyglm object
severeMDEMODEL2015to2019LG_OR
```
### Bisexual 
```{r}
severeMDEMODEL2015to2019Bi <- survey::svyglm(amdeimp ~ EcstasyVariable + PSILCY2+ lsdflag + PEYOTE2 +MESC2 + MarriageVariable+ EducationVariable+ CATAG6+ NEWRACE2+ irsex+ income+
                                           RiskyVariable+ cocflag+ herflag+ pcpflag+ InhalantVariable + PainRelieverVariable+ TranquilizerVariable+KetamineVar + PoppersVariable+ 
                                           StimulantVariable+ SedativeVariable+ mrjflag+year, data=NSDUH2015to2019DataBis,family=quasibinomial(), design=biDesign)
summary(severeMDEMODEL2015to2019Bi)
severeMDEMODEL2015to2019Bi_OR <- or_table(severeMDEMODEL2015to2019Bi)   # display only; keeps severeMDEMODEL2015to2019Bi as the svyglm object
severeMDEMODEL2015to2019Bi_OR
```

# ============================================================
# Tables for publication
# ============================================================

## Table 1 — Sample demographics, psychedelic use, mental-health outcomes, and covariates by sexual orientation
Survey-weighted percentages; categorical comparisons use svychisq (Rao–Scott second-order correction).
```{r}

# tableone::svyCreateTableOne builds a survey-weighted "Table 1" and tests categorical
# variables with svychisq(), whose default F statistic is the Rao-Scott second-order correction.
t1_vars <- c(
  "EcstasyVariable", "PSILCY2",                                   # psychedelic use
  "irsex", "CATAG6", "NEWRACE2", "EducationVariable", "income",
  "MarriageVariable", "RiskyVariable",                            # demographics
  "lsdflag", "PEYOTE2", "MESC2", "cocflag", "herflag", "pcpflag",
  "InhalantVariable", "PainRelieverVariable", "TranquilizerVariable",
  "StimulantVariable", "SedativeVariable", "mrjflag",
  "KetamineVar", "PoppersVariable",                              # other lifetime substance use
  "spdmon", "mhsuithk", "mhsuipln", "amdelt", "amdeyr", "amdeimp" # mental-health outcomes
)

table1 <- svyCreateTableOne(
  vars       = t1_vars,
  strata     = "sexident",
  data       = totalDesign,     # survey design with pooled 5-year weights
  factorVars = t1_vars,         # treat all variables as categorical -> weighted percentages
  test       = TRUE             # svychisq() default = Rao-Scott second-order F
)

# Capture the formatted table as a character matrix WITHOUT console-printing it, then render a
# clean table with knitr::kable. (Printing the raw tableone object is what looks "weird" --
# it prints a quoted, misaligned character matrix.)
t1mat <- print(
  table1,
  format        = "p",     # survey-weighted percentages
  digits        = 1,
  pDigits       = 5,
  explain       = FALSE,
  showAllLevels = FALSE,   # one row per binary variable; set TRUE to show every level
  noSpaces      = TRUE,    # strip leading spaces that misalign columns
  printToggle   = FALSE
)

knitr::kable(
  t1mat,
  caption = "Table 1. Sample demographics, psychedelic use, mental-health outcomes, and covariates by sexual orientation (survey-weighted %; Rao-Scott second-order chi-square)."
)

# Export:
# write.csv(t1mat, "table1_demographics.csv")
```

## Table 2 — interaction screen (sexual identity x drug)
```{r}
interaction_models <- c(
  "SPD (PM) x Psilocybin"               = "PSYDISTMODEL2015to2019INTERpsilthree",
  "Suicidal Ideation (PY) x Psilocybin" = "SUITHINKMODEL2015to2019INTERpsilthree",
  "Suicidal Ideation (PY) x MDMA"       = "SUITHINKMODEL2015to2019INTERecstthree",
  "Suicidal Planning (PY) x MDMA"       = "SUICPLANMODEL2015to2019INTERecstthree",
  "Past-Year MDE x Psilocybin"          = "pstyrMDEMODEL2015to2019INTERpsilthree",
  "Past-Year MDE x MDMA"                = "pstyrMDEMODEL2015to2019INTERecstthree",
  "Lifetime MDE x Psilocybin"           = "lifetimeMDEMODEL2015to2019INTERpsilthree",
  "Lifetime MDE x MDMA"                 = "lifetimeMDEMODEL2015to2019INTERecstthree",
  "Severe MDE (PY) x Psilocybin"        = "severeMDEMODEL2015to2019INTERpsilthree",
  "Severe MDE (PY) x MDMA"              = "severeMDEMODEL2015to2019INTERecstthree"

)

table2 <- do.call(rbind, lapply(names(interaction_models), function(lab){
  obj <- interaction_models[[lab]]
  if (!exists(obj) || !inherits(get(obj), "svyglm")) {
    warning("Skipping (missing/overwritten): ", obj); return(NULL)
  }
  tab <- or_table(get(obj))
  int <- tab[grepl(":", tab$term), ]                 # sexident x drug interaction rows
  if (nrow(int) == 0) return(NULL)
  data.frame(screen = lab, int, row.names = NULL)
}))
table2$CI   <- sprintf("%.2f (%.2f, %.2f)", table2$OR, table2$lower, table2$upper)
table2$q_BH <- round(p.adjust(table2$p, method = "BH"), 5)   # FDR across the screen
# The p-value is the quantity of interest for the moderation screen; the interaction OR is
# reported for completeness only. Flag significant rows (*) and foreground p (smallest first).
table2$sig  <- ifelse(table2$p <= .05, "*", "")
table2 <- table2[order(table2$p), c("screen", "term", "p", "sig", "q_BH", "OR", "CI")]
table2$p    <- formatC(table2$p,    format = "f", digits = 5)   # fixed 5 dp: borderline reads e.g. 0.04950, not 0.05
table2$q_BH <- formatC(table2$q_BH, format = "f", digits = 5)
print(table2, row.names = FALSE)
# Optional visually highlighted version for the knitted supplement (requires kableExtra):
# library(kableExtra)
# kbl(table2, caption = "Interaction screen (p-value is the quantity of interest)") |>
#   kable_styling() |> column_spec(3, bold = TRUE) |>
#   row_spec(which(table2$sig == "*"), background = "#FFF3CD")
# write.csv(table2, "table2_interaction_screen.csv", row.names = FALSE)
```

## Table 3 — stratified subgroup models (psilocybin main effect within identity)
```{r}
strat_spec <- list(
  list(lab="Severe MDE (PY) | Heterosexual",        obj="severeMDEMODEL2015to2019HET"),
  list(lab="Severe MDE (PY) | Gay/Lesbian",         obj="severeMDEMODEL2015to2019LG"),
  list(lab="Severe MDE (PY) | Bisexual",            obj="severeMDEMODEL2015to2019Bi"),
  list(lab="Suicidal Ideation (PY) | Heterosexual", obj="SUITHINKMODEL2015to2019HET"),
  list(lab="Suicidal Ideation (PY) | Gay/Lesbian",  obj="SUITHINKMODEL2015to2019Lez"),
  list(lab="Suicidal Ideation (PY) | Bisexual",     obj="SUITHINKMODEL2015to2019Bi")
)
table3 <- do.call(rbind, lapply(strat_spec, function(s){
  if (!exists(s$obj) || !inherits(get(s$obj), "svyglm")) {
    warning("Skipping (missing/overwritten): ", s$obj); return(NULL)
  }
  tab <- or_table(get(s$obj))
  row <- tab[grepl("^PSILCY2", tab$term), ]          # psilocybin main effect in stratum
  if (nrow(row) == 0) return(NULL)
  data.frame(stratum = s$lab, row, row.names = NULL)
}))
table3$CI <- sprintf("%.2f (%.2f, %.2f)", table3$OR, table3$lower, table3$upper)
table3 <- table3[, c("stratum", "CI", "p")]
table3$p <- formatC(table3$p, format = "f", digits = 5)   # fixed 5 dp
print(table3)
# write.csv(table3, "table3_stratified.csv", row.names = FALSE)
```

## Consistency check — every interaction model's CIs and p-values should agree
```{r}
invisible(lapply(interaction_models, function(obj){
  if (exists(obj) && inherits(get(obj), "svyglm")) { cat(obj, ": "); check_consistency(get(obj)) }
}))
```


## Supplementary Table — full-sample associations (psilocybin & MDMA/ecstasy main effects, all respondents)

```{r}
fullsample_models <- c(
  "Past-month psychological distress" = "PSYDISTMODEL2015to2019",
  "Past-year suicidal ideation"       = "SUITHINKMODEL2015to2019",
  "Past-year suicidal planning"       = "SUIPLANMODEL2015to2019",
  "Past-year MDE"                     = "pstyrMDEMODEL2015to2019",
  "Lifetime MDE"                      = "lifetimeMDEMODEL2015to2019",
  "Severe past-year MDE"              = "severeMDEMODEL2015to2019"
)

full_sample_table <- do.call(rbind, lapply(names(fullsample_models), function(lab){
  obj <- fullsample_models[[lab]]
  if (!exists(obj) || !inherits(get(obj), "svyglm")) { warning("Missing model: ", obj); return(NULL) }
  tab  <- or_table(get(obj))                                  # design-df OR / 95% CI / p
  rows <- tab[grepl("^PSILCY2", tab$term) | grepl("^EcstasyVariable", tab$term), ]
  if (nrow(rows) == 0) return(NULL)
  rows$Substance <- ifelse(grepl("^PSILCY2", rows$term), "Psilocybin", "MDMA/ecstasy")
  rows$aOR_95CI  <- sprintf("%.2f (%.2f, %.2f)", rows$OR, rows$lower, rows$upper)
  data.frame(Outcome = lab, Substance = rows$Substance,
             aOR = rows$OR, aOR_95CI = rows$aOR_95CI, p = rows$p, row.names = NULL)
}))
# order psilocybin then MDMA within each outcome, in the outcome order above
full_sample_table <- full_sample_table[
  order(match(full_sample_table$Outcome, names(fullsample_models)),
        full_sample_table$Substance == "MDMA/ecstasy"), ]
full_sample_table$p <- formatC(full_sample_table$p, format = "f", digits = 5)   # fixed 5 dp
print(full_sample_table, row.names = FALSE)
# write.csv(full_sample_table, "table_full_sample_associations.csv", row.names = FALSE)
```

## Session information (for reproducibility)
```{r}
sessionInfo()
```
